---
title: "Zach NDMS"
author: "Daniel J. Naumenko"
date: "July 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Zach notes:
Alright so this is 16s data from a series of "incubation vessels". The water in those were either filtered or unfiltered and the species of Crustose Coraline Algae inside were also different. So I am essentially trying to figure out what the effect of the treatments are on the communities.

Good luck.

```{r Libraries, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# import libraries

library(dplyr)
library(ggplot2)
library(readr)
library(MASS)
library(vegan)
```

I made minor manual edits to the column names in the original csv file. It was easier to add underscores in spaces between some column names (and in one case replace parentheses with underscores) to avoid issues with standard R naming conventions.
```{r Data import, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# read in data
microb_NT <- read.csv("CRUSADE_OTUS_Master_Cleaned.csv")

#transform data arc(sqrt())
microb_df <- microb_NT%>%
  mutate_if(is.numeric, sqrt)%>%
  mutate_if(is.numeric, asin)

# edit column names to contain no spaces
microb_abun <- microb_df %>% 
  mutate()%>%
  dplyr::filter(Timepoint == "T3") %>% 
  dplyr::filter(!Organism == "Water Control")%>%
  dplyr::select(`Label`:`Verrucomicrobiales_Verrucomicrobiaceae_Roseibacillus.sp`)

# remove original dataframe
rm(microb_df)

```

Ok, so the only columns that are required for NDMS calculations are Alteromonadales_Alteromonadaceae_Alteromonas_sp:Flavobacteriales_Flavobacteriaceae_Cytophaga_sp_41-DBG2_sp. I am following methods of Jari Oksanen in "Multivariate Analysis of Ecological Communities in R: vegal tutorial. June 10, 2015", secion 2.1: NDMS

NDMS can be performed with isoMDS(), but requires dissimilarities as input. Bray-Curtis dissimilarity is the default (also known as Steinhaus dissimularity, or the Sorensen index in Finland).


```{r eval=FALSE, fig.cap="The stressplot function draws a Shepard plot where ordination distances are plotted against community dissimilarities (using Bray-Curtis dissimilarity as default, see above). Also shows two correlation-like goodness of fit statistics. The non-metric fit is based on Stress (1-stress^2) where Stress is a function of and non-linear monotone transformation of observed dissimilarities and ordination distances. The fit-based Rsq: is correlation between the fitted values and ordination distances.", message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

stressplot(vare.mds0, vare.dis)
```




```{r}


# create community x species matrix
microb <- as.matrix(microb_abun %>% 
                      dplyr::select(`Subgroup_3_PAUC26f`:`Verrucomicrobiales_Verrucomicrobiaceae_Roseibacillus.sp`),
                          dimnames = list(paste("Sample", 1:31, sep = ",")))

# Run NmDS, provide community x species matrix, and set dimensions to 2
# outputs each iteration until a solution is reached (minimize stress after reconfiguration of points in 2D)

set.seed(1234)
microb_NMDS <- metaMDS(microb, k = 2)

# produce Shepard plot
# large scatter around the line suggests that original dissimilarities are not well preserved in the reduced number of dimensions. These look pretty good!
stressplot(microb_NMDS)

# plot the NMDS
# open circles are communities, red crosses are species
plot(microb_NMDS)

ordiplot(microb_NMDS, type = "n")
orditorp(microb_NMDS, display = "sites", air = 0.25)



# create a plot with convex hulls connecting vertices of the points made by a treatment

treatment <- c(microb_abun$Water.Treatment)
samples <- c(microb_abun$Label)

ordiplot(microb_NMDS, type = "n",
         main = "NMDS of Incubation Vessels")
ordihull(microb_NMDS, groups = treatment, draw = "polygon", label = F)
#orditorp(microb_NMDS, display = "sites", labels = samples, air = 0.05)
orditorp(microb_NMDS, display = "sites", air = 0.2)


# spider plot
ordiplot(microb_NMDS, type = "n",
         main = "NMDS of Incubation Vessels")
orditorp(microb_NMDS, display = "sites", air = 0.2)
ordispider(microb_NMDS, groups = treatment)


# ellipse plot
ordiplot(microb_NMDS, type = "n",
         main = "NMDS of Incubation Vessels")
orditorp(microb_NMDS, display = "sites", air = 0.2)
ordiellipse(microb_NMDS, groups = treatment)



# MST plot

dist_microb <- vegdist(microb)
clust_microb <- hclust(dist_microb, method = "complete")


ordiplot(microb_NMDS, type = "n",
         main = "NMDS of Incubation Vessels")
orditorp(microb_NMDS, display = "sites", air = 0.2)
ordicluster(microb_NMDS, cluster = clust_microb)



```


```{r}
# plot NMDS output in ggplot

df.scores <- as.data.frame(scores(microb_NMDS)) #Using the scores function from vegan to extract the site scores and convert to a data.frame

df.scores$site <- rownames(df.scores) # create a column of site names, from the rownames of data.scores

# add grouping variables
df.scores$sample <- microb_abun$Label
df.scores$group <- microb_abun$Water.Treatment
df.scores$dendolabel <- microb_abun$Dendolabel
df.scores$DNA_source <- microb_abun$DNA.Source
df.scores$Organism <- microb_abun$Organism


species.scores <- as.data.frame(scores(microb_NMDS, "species")) #Using the scores function from vegan to extract the species scores and convert to a data.frame

species.scores$species <- rownames(species.scores)



# plot NMDS

ggplot() +
  geom_point(data = df.scores, aes(x = NMDS1, y = NMDS2, shape = group), alpha = 0.5) +
  geom_text(data = df.scores, aes(x = NMDS1, y = NMDS2, label = site)) +
  coord_equal() +
  scale_shape_manual(values = c(0, 19)) +
  theme_bw()



# plot with convex group hulls

grp.filter <- df.scores[df.scores$group == "Filtered",][chull(df.scores[df.scores$group == "Filtered", c("NMDS1", "NMDS2")]), ] # hull values for filtered group

grp.unfilter <- df.scores[df.scores$group == "Unfiltered",][chull(df.scores[df.scores$group == "Unfiltered", c("NMDS1", "NMDS2")]), ] # hull values for unfiltered group

hull.data <- rbind(grp.filter, grp.unfilter)


ggplot() +
  geom_polygon(data = hull.data, aes(x = NMDS1, y = NMDS2, fill = group, group = group), alpha = 0.3) +
  geom_text(data = species.scores, aes(x = NMDS1, y = NMDS2, label = species),
            alpha = 0.5) +
  geom_point(data = df.scores, aes(x = NMDS1, y = NMDS2, shape = group, colour = group), size = 3) +
  geom_text(data = df.scores, aes(x = NMDS1, y = NMDS2, label = site)) +
  coord_equal() +
  theme_bw()
```



```{r eval=FALSE, include=FALSE}

# write csv for Zach to plot in JMP
NMDS_scores <- as.data.frame(scores(microb_NMDS, display = c("sites")))


# add grouping variables
NMDS_scores$Sample <- microb_abun$Label
NMDS_scores$Water_Treatment <- microb_abun$Water.Treatment
NMDS_scores$DNA_source <- microb_abun$DNA.Source
NMDS_scores$Organism <- microb_abun$Organism



write.csv(NMDS_scores, "NMDS_scores.csv")
```

